---
title: "External validation"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r}
# create this vignette
# devtools::use_vignette("external-validation")
```


External validation, or model evaluation, is the process of comparing the output
of spatial microsimulation models with data is *external* to the model. It
is often difficult to do external validation because one of the main reasons
for generating synthetic spatial microdata is that real microdata does
not exist. So there may be nothing to compare the data against.

To overcome this issue, we use public spatial microdata data from the UK's Office
of National Statistics (ONS). This is real data, a 1% sample of anonymised
records 




```{r, eval=FALSE}
# Warning: this will download 7 MB of data, may take some time
hlink <- "http://www.ons.gov.uk/ons/rel/census/2011-census/2011-census-teaching-file/rft-teaching-file.zip"
download.file(hlink, destfile = "rft-teaching-file.zip")
dir.create("external-validation")
# Extract the data to a new folder
unzip("rft-teaching-file.zip", exdir = "external-validation")
```

```{r, eval=FALSE}
# smdata <- read.csv("../external-validation/2011 Census Microdata Teaching File.csv", skip = 1)
smdata <- read.csv("external-validation/2011 Census Microdata Teaching File.csv", skip = 1)
head(smdata)
summary(smdata$Region) # the spatial data that we will seek to estimate
```

## Generating the sample microdata

With almost 600,000 rows, the spatial microdata loaded is sufficiently large
to sample, whilst still retaining good distribtion of individual-level variables.

```{r, eval=FALSE}
sel <- sample(nrow(smdata), size = nrow(smdata) / 100)
ind <- smdata[sel, ] # create microdata
ind$Region <- NULL # remove geographical information
```





## Vignette info

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style


Note the various macros within the `vignette` setion of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
