---
title: "testmsim: a package for testing spatial microsimulation methods"
author: "Esteban Munoz and Robin Lovelace"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo=FALSE}
# This is a description of the testsmsim package, it motivations and some results
# it has generated so far
# location: vignettes/testmsim-paper.Rmd

# Libraries needed to compile this document
libs <- c("knitr", "devtools")
lapply(libs, library, character.only = T)

# Installing the testmsim R library (uncomment to install latest version)
# install_github("robinlovelace/testmsim")
```

## Introduction

'Spatial microsimulation' is a term that has become associated with methods
of synthesising and analyzing multi-level data: individuals (typically humans,
households, firms or ...) allocated to geographical space, typically
administrative zones. In the terminology of the spatial microsimulation
literature, this type of individual-level data allocated to zones is
called 'spatial microdata' [@ref1; @ref2]. Spatial microdata is usually
stored as a long 2 dimensional table resembling the general structure
illustrated in Table 1. Note that the data could equally be stored as
a `list` type, with each zone represented by a separate list object. However,
for analysis purposes, the 'long data format presented in Table 1 is
generally preferable as it is 'tidy' [@tidy-data].


```{r, echo=FALSE, eval=FALSE}
df1 <- read.csv(textConnection("ID,V1,V2,V...,Vn,Zone
1,,,,,1
2,,,,,1
3,,,,,1
…,,,,,…
…,,,,,2
…,,,,,2
…,,,,,…
T,,,,,Zm"), )
df1[2:5] <- " "
kable(df1, row.names = F)
```

|ID |V1 |V2 |V... |Vn |Zone |
|:--|:--|:--|:----|:--|:----|
|1  |   |   |     |   |1    |
|2  |   |   |     |   |1    |
|3  |   |   |     |   |1    |
|…  |   |   |     |   |…    |
|…  |   |   |     |   |2    |
|…  |   |   |     |   |2    |
|…  |   |   |     |   |…    |
|T  |   |   |     |   |Zm   |


There is an ongoing debate about whether the term 'spatial microsimulation'
should be used to describe the *process* of generating such spatial microdata
[ref] or an *approach* to investigating multi-level problems that uses spatial
microdata as its foundation [ref].
the process of generating 

## Data

## Methods

The methods presented in this paper are threefold: 

1. Code (saved in functions) and test data for generating spatial microdata from 
	aggregate and individual-level data using a variety of techniques.
2. A collection of test functions for internal validation of microsimulation models.
3. A collection of methods for the external validation of these models.

The first set of functions consists of 13 test. We classify these test into two
categories: a) tests comparing design survey weights with estimated new weights;
and b) tests comparing small area census data with the marginal sums of the 
resulting synthetic population.

### Methods of population synthesis

#### IPF

#### GREGWT

```{r, echo=FALSE}
# I think we should just start with these two. We can add more later (RL)
```

### Weight based tests

**Weights Distance**

Computes the distance between sample design weights and estimated new
weights.

$ D_i = \sum_j^m |w_j - d_j| $

**Total Chi-squared distance**

Computes the Chi-squared distance between sample design weights and
estimated new weights.

$ Chi_i = \sum_j^m \frac{\left( w_j \times d_j\right)^2 }{2d_j} $

**Mean Chi-squared distance**

Computes the mean Chi-squared distance between sample design weights and
estimated new weights.

$ \theta Chi_i = \sum_j^m \frac{\left(w_j \times d_j\right)^2 }{2d_j} \div m $

**Total absolute distance (TAD)**

Computes the total absolute distance between sample design weights and
estimated new weights.

$ TAD = \sum_i^n \left|\sum_j^m w_{i,j} - pop_i\right| $

**Error in Margin (EM)**

Computes the ration between estimated and known number of individuals or
households.

$ EM_i = \frac{\sum w_j - pop_i}{pop_i} $

**Error in Distribution (ED)**

Computes the ration between absolute sum of residuals (estimated - known)
and the actual population size.

$ ED_i = \frac{|\sum w_j - pop_i|}{pop_i} $

### Marginal sums based tests

**Total absolute error (TAE)**

Computes the total absolute error as the sum of the absolute difference
between observed marginal sums and simulated marginal sums.

$TAE = \sum_i^n |Tx - \hat{t}x|$

**Standardized absolute error (SAE)**

Divides the \code{\link{getTAE}} by the know population size.

$ SAE = \sum_i^n |Tx - \hat{t}x| \div pop_i $

**Percentage error (PSAE)**

Divides the \code{\link{getSAE}} by the known population size.

$ PAE = \sum_i^n |Tx - \hat{t}x| \div pop_i \times 100 $

**Z-statistic**

The Z-statistic aims to describe the performance of the individual
characteristics of the population used as constrains in the simulation.

$$
   r = \frac{\hat{t}x}{\sum Tx}
   p = \frac{Tx}{\sum Tx}
   Z = \frac{r-p}{\sqrt{p\times\left(1-p\right)\div\sum Tx}}
$$

**Correlation Coefficient (Pearson Correlation)**

test for correlations between simulated and expected marginal totals. This
functions implements the \code{\link[stats]{cor}} function for the
estimation of the Pearson correlation coefficient.

``pearson <- cor(cbind(Tx, hTx), use="complete.obs", method="pearson")``

**Independent samples t-Test**

performs a t-test to compare expected proportions between simulated and
observed marginal totals. This function implements the
available `t.test` function in R.

**Coefficient of determination**

Compute the r-squared coefficient of determination between simulated and
observed marginal totals. This function implements the function `lm` in R 
to estimate the r-squared coefficient.

## Results

## Discussion